---
title: "00 Setup Data Symlinks & Cryptographic Pointers"
author: "Agustin Carbajal & Research Team"
date: today
format:
  html:
    toc: true
    code-fold: show
# In-Document Provenance:
# R: 4.3+ | tidyverse: 2.0+ | fs: 1.6+ | data_tracker.sh integration
---

## 1. Introduction & Workflow

This Quarto document replaces manual, unversioned symlink scripts (such as legacy `00_Symlinks.Rmd`).
Instead of fragile `../../../` relative paths that break across machines, this workflow:
1. Resolves external datasets portably via `$DATA_ROOT` (defined in `.env`).
2. Registers cryptographic hashes into `local_pointers.tsv`.
3. Creates organized local symlinks in `raw_data/` (e.g., `raw_data/Resources/`, `raw_data/sciATAC_fragment_files/`).
4. Verifies data integrity with the Stale Checksum Guard before any analysis begins.

---

## 2. Environment & Helper Functions

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(fs)

# Load DATA_ROOT from project .env or system environment
get_data_root <- function() {
  # Check if .env exists in project root
  if (file.exists(".env")) {
    readRenviron(".env")
  }
  
  root <- Sys.getenv("DATA_ROOT", unset = "")
  if (root == "" || !dir.exists(root)) {
    stop("DATA_ROOT is not set or directory does not exist. Please check your .env file.")
  }
  return(root)
}

data_root <- get_data_root()
message("Using external DATA_ROOT: ", data_root)

# Helper function to track & link a single file using data_tracker.sh
track_dataset <- function(link_name, rel_source_path, rel_checksum_path) {
  res <- system2("./scripts/data_tracker.sh", 
                 args = c("add", link_name, rel_source_path, rel_checksum_path),
                 stdout = TRUE, stderr = TRUE)
  status <- attr(res, "status")
  if (!is.null(status) && status != 0) {
    stop("Failed to track: ", link_name, "\n", paste(res, collapse = "\n"))
  }
  message("✓ Tracked & linked: ", link_name)
}
```

---

## 3. Provision Reference Resources

Link individual reference files into organized subdirectories inside `raw_data/Resources/`:

```{r resources}
# Upstream references often have their own checksums file (e.g., in References/Mice/checksums.tsv)
# track_dataset(
#   link_name = "Resources/tes.expanded.bed",
#   rel_source_path = "References/Mice/cell_ranger_datasets/refdata-cellranger-arc-mm10-2020-A-2.0.0/tes.expanded.bed",
#   rel_checksum_path = "References/Mice/checksums.tsv"
# )

# track_dataset(
#   link_name = "Resources/tss.expanded.bed",
#   rel_source_path = "References/Mice/cell_ranger_datasets/refdata-cellranger-arc-mm10-2020-A-2.0.0/tss.expanded.bed",
#   rel_checksum_path = "References/Mice/checksums.tsv"
# )
```

---

## 4. Batch Discovery & Linking (sciATAC Fragment Files)

Using R's pattern matching (`list.files`) to batch-register multiple fragment files from a cohort,
referencing a higher-level upstream `checksums.tsv` file:

```{r batch_fragments}
# Define the cohort folder relative to DATA_ROOT
cohort_dir <- "sciATAC-seq/sciATAC_seq_Sep_2024_HP1ag_dKO_CHD4_KO_BAF200_KO/fragment_files"
abs_cohort_dir <- file.path(data_root, cohort_dir)

# Higher-level checksum file residing one level up (or at the root of the cohort)
rel_checksum_file <- "sciATAC-seq/sciATAC_seq_Sep_2024_HP1ag_dKO_CHD4_KO_BAF200_KO/checksums.tsv"

if (dir.exists(abs_cohort_dir)) {
  # Discover files using regex pattern matching
  frag_files <- list.files(
    path = abs_cohort_dir,
    pattern = "CHD4|WT.*\\.gz$",
    full.names = FALSE
  )
  
  message(sprintf("Discovered %d matching fragment files in cohort.", length(frag_files)))

  # Batch register all discovered files
  walk(frag_files, function(f) {
    track_dataset(
      link_name = file.path("sciATAC_fragment_files", f),
      rel_source_path = file.path(cohort_dir, f),
      rel_checksum_path = rel_checksum_file
    )
  })
} else {
  message("Cohort directory not found on current DATA_ROOT mount. (Example template)")
}
```

---

## 5. Verify All Pointers

Execute Tier 1 verification to ensure all local symlinks are healthy and hashes match:

```{r verify}
verify_output <- system("./scripts/data_tracker.sh verify", intern = TRUE)
cat(verify_output, sep = "\n")
```

---

## 6. Session Information & Provenance

```{r session_info}
sessioninfo::session_info()
```
