---
title: "Genomics Analysis (R Template)"
author: "Research Team"
date: today
format:
  html:
    toc: true
    code-fold: show
# In-Document Provenance: Core tool versions tested for this analysis
# R: 4.3+ | BiocManager: 3.18+ | DESeq2: 1.42+ | tidyverse: 2.0+
---

## 1. Introduction & Objectives

This notebook demonstrates the in-document version-locking pattern for R-based Quarto analyses.
Input raw data is accessed via symlinks in `raw_data/` provisioned by `scripts/data_tracker.sh`.

## 2. Load Libraries

```{r}
#| message: false
#| warning: false
library(tidyverse)
# library(DESeq2)
# library(GenomicRanges)
```

## 3. Data Ingestion via Locked Pointers

```{r}
# Data files are accessed through the gitignored raw_data/ symlink directory
data_file <- "raw_data/sample1.bam"
if (file.exists(data_file)) {
  message("Found locked input data: ", data_file)
} else {
  message("Raw data symlink not yet provisioned. Run: ./scripts/data_tracker.sh link")
}
```

## 4. Analysis Workflow

```{r}
# Perform reproducible analysis steps here
# df <- read_tsv(...)
```

## 5. Session Information & In-Document Version Locking

> [!NOTE]
> Rendering `sessionInfo()` in the final chunk permanently records the exact R version,
> OS platform, and attached package versions into the Git-committed `.qmd` and rendered report.

```{r}
#| echo: true
sessioninfo::session_info()
```
